Models

### 2022-04-13 analysis/05_model_metrics.Rmd
Fit models for all CWM models, add scenarios where there aren't enough data to model fit function

model_data <- metrics %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "Jaccard_next",
                       "FRic", "FEve", "FDiv", "S",
                       "bodymass_value"),
         !rarefyID %in% missing_null) %>%
  group_by(metric) %>%
  mutate(study_id =  str_extract(rarefyID, "[^_]+")) %>%
  mutate(year_scaled = scale(year), 
         scale_center = attributes(year_scaled)$`scaled:center`, 
         scale = attributes(year_scaled)$`scaled:scale`) %>%
  ungroup() %>%
  #this study has many low species richness observations, so FD metrics weren't calculated, just drop it
  filter(study_id != 348)

@@

@@

  if(!model_id %in% c("base", "base_dur", "base_srt")){
    slopes <- get_fixed_slopes(metric = metric_name, model = fit$result, model_id = model_id, model_data = fit_data)

@@

    df <- tidy(fit$result) %>% 
      bind_rows(group_slopes, slope_constrasts) %>%
      mutate(metric = metric_name, model = list(fit$result), model_id = model_id, warning = ifelse(length(fit$warnings) == 0, NA, fit$warnings))
  }else{
  tidy(fit$result) %>%
   mutate(metric = metric_name, model = list(fit$result), model_id = model_id, warning = ifelse(length(fit$warnings) == 0, NA, fit$warnings))
  }
}

@@

Create dataframe of models to fit
```{r}
raw_value_mods <- tibble(metric_name = c("SES_FRic", "SES_FEve", "SES_FDiv", "FRic", "FEve", "FDiv", "Jaccard_base"), 
                         base = "value ~ year_scaled + (year_scaled|study_id/rarefyID)",
                         taxa = "value ~ year_scaled * taxa + (year_scaled|study_id/rarefyID)",
                         realm = "value ~ year_scaled * realm + (year_scaled|study_id/rarefyID)",
                         climate = "value ~ year_scaled * climate + (year_scaled|study_id/rarefyID)") %>%
  pivot_longer(cols = c(base, taxa, realm, climate), names_to = "model_id", values_to = "formula")

@@

Models with interactions
```{r}
int_raw_value_mods <- tibble(metric_name = c("SES_FRic", "SES_FEve", "SES_FDiv", "FRic", "FEve", "FDiv", "Jaccard_base"), 
                         base_dur = "value ~ year_scaled + duration_scaled*year_scaled + (year_scaled|study_id/rarefyID)",
                         base_srt = "value ~ year_scaled + startyear_scaled*year_scaled + (year_scaled|study_id/rarefyID)") %>%
  pivot_longer(cols = c(base_dur, taxa_dur, realm_dur, climate_dur, base_srt, taxa_srt, realm_srt, climate_srt), 
               names_to = "model_id", values_to = "formula")

int_log_value_mods <- tibble(metric_name = c("S"),
                         base_dur = "logvalue ~ year_scaled + duration_scaled*year_scaled + (year_scaled|study_id/rarefyID)",
                         base_srt = "logvalue ~ year_scaled + startyear_scaled*year_scaled + (year_scaled|study_id/rarefyID)") %>%
  pivot_longer(cols = c(base_dur, taxa_dur, realm_dur, climate_dur, base_srt, taxa_srt, realm_srt, climate_srt), 
               names_to = "model_id", values_to = "formula")

# mods <- bind_rows(raw_value_mods, log_value_mods) %>% 
#   select(metric_name, formula, model_id)

#Fit models with base settings
int_mod_results <- pmap_dfr(bind_rows(int_raw_value_mods, int_log_value_mods), get_model, data = model_data)

#get the models that had convergence warnings
bad_models <- int_mod_results %>%
  select(metric, model_id, warning) %>% 
  filter(!is.na(warning)) %>% 
  distinct() %>%
  left_join(int_raw_value_mods, by = c("metric" = "metric_name", "model_id")) 
```

>>>>

model_data <- metrics %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "Jaccard_next",
                       "FRic", "FEve", "FDiv", "S","CWM_bodymass_value","CWM_diet_inv", 
                       "CWM_diet_scav", "CWM_diet_vfish", "CWM_diet_vunk", "CWM_forstrat_ground", 
                       "CWM_forstrat_wataroundsurf", "CWM_forstrat_watbelowsurf", "CWM_diet_fruit", 
                       "CWM_diet_planto", "CWM_diet_seed", "CWM_diet_vend", "CWM_forstrat_understory", 
                       "CWM_diet_vect", "CWM_diet_nect", "CWM_forstrat_aerial", "CWM_forstrat_canopy", 
                       "CWM_forstrat_midhigh", "CWM_age_at_maturity_max_y", "CWM_age_at_maturity_min_y", 
                       "CWM_body_size_mm", "CWM_litter_size_max_n", "CWM_litter_size_min_n", 
                       "CWM_longevity_max_y", "CWM_offspring_size_max_mm", "CWM_offspring_size_min_mm"),
         !rarefyID %in% missing_null) %>%
  group_by(metric) %>%
  mutate(study_id =  str_extract(rarefyID, "[^_]+")) %>%
  mutate(year_scaled = scale(year), 
         scale_center = attributes(year_scaled)$`scaled:center`, 
         scale = attributes(year_scaled)$`scaled:scale`) %>%
  ungroup() %>%
  #this study has many low species richness observations, so FD metrics weren't calculated, just drop it
  filter(study_id != 348)

@@

  if(n_distinct(fit_data$rarefyID) < 2){
    return(tibble())
  }

  if(!model_id %in% c("base", "base_dur", "base_srt")){
    if(fit_data %>% pull(model_id) %>% n_distinct() >1) {

@@

      slopes <- get_fixed_slopes(metric = metric_name, model = fit$result, model_id = model_id, model_data = fit_data)

@@

      df <- tidy(fit$result) %>% 
        bind_rows(group_slopes, slope_constrasts) %>%
        mutate(metric = metric_name, model = list(fit$result), model_id = model_id,
               n_ts = n_distinct(fit_data$rarefyID),
               warning = ifelse(length(fit$warnings) == 0, NA, fit$warnings))
    }else{
      return(tibble())
    }
  }else{

    tidy(fit$result) %>%
      mutate(metric = metric_name, model = list(fit$result), model_id = model_id, 
             n_ts = n_distinct(fit_data$rarefyID),
             warning = ifelse(length(fit$warnings) == 0, NA, fit$warnings))
  }
}

@@

Create dataframe of models to fit
```{r}
metric_names <- unique(model_data$metric)
raw_value_mods <- tibble(metrimod_resultsc_name = metric_names[metric_names != "S"], 
                         base = "value ~ year_scaled + (year_scaled|study_id/rarefyID)",
                         taxa = "value ~ year_scaled * taxa + (year_scaled|study_id/rarefyID)",
                         realm = "value ~ year_scaled * realm + (year_scaled|study_id/rarefyID)",
                         climate = "value ~ year_scaled * climate + (year_scaled|study_id/rarefyID)") %>%
  pivot_longer(cols = c(base, taxa, realm, climate), names_to = "model_id", values_to = "formula")

@@

Write out table of model estimates for manuscript
```{r}
metric_model_table %>% 
  filter(model_id == "base", metric %in% c("S", "Jaccard_base", "SES_FRic", "SES_FEve", "SES_FDiv")) %>% 
  select(-c(model, model_id, warning, statistic, df)) %>% 
  write_csv(here::here("figures", "man_model_table.csv"))
```

Models with interactions
```{r}
# int_raw_value_mods <- tibble(metric_name = c("SES_FRic", "SES_FEve", "SES_FDiv", "FRic", "FEve", "FDiv", "Jaccard_base"), 
#                          base_dur = "value ~ year_scaled + duration_scaled*year_scaled + (year_scaled|study_id/rarefyID)",
#                          base_srt = "value ~ year_scaled + startyear_scaled*year_scaled + (year_scaled|study_id/rarefyID)") %>%
#   pivot_longer(cols = c(base_dur, taxa_dur, realm_dur, climate_dur, base_srt, taxa_srt, realm_srt, climate_srt), 
#                names_to = "model_id", values_to = "formula")
# 
# int_log_value_mods <- tibble(metric_name = c("S"),
#                          base_dur = "logvalue ~ year_scaled + duration_scaled*year_scaled + (year_scaled|study_id/rarefyID)",
#                          base_srt = "logvalue ~ year_scaled + startyear_scaled*year_scaled + (year_scaled|study_id/rarefyID)") %>%
#   pivot_longer(cols = c(base_dur, taxa_dur, realm_dur, climate_dur, base_srt, taxa_srt, realm_srt, climate_srt), 
#                names_to = "model_id", values_to = "formula")
# 
# 
# # mods <- bind_rows(raw_value_mods, log_value_mods) %>% 
# #   select(metric_name, formula, model_id)
# 
# #Fit models with base settings
# int_mod_results <- pmap_dfr(bind_rows(int_raw_value_mods, int_log_value_mods), get_model, data = model_data)
# 
# #get the models that had convergence warnings
# bad_models <- int_mod_results %>%
#   select(metric, model_id, warning) %>% 
#   filter(!is.na(warning)) %>% 
#   distinct() %>%
#   left_join(int_raw_value_mods, by = c("metric" = "metric_name", "model_id")) 
```
